\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-19 notes/2loopNLSM.tex
Conclusion + misc

We can achieve this construction by splitting the gauge group of the full Yang-Mills action from $G \rightarrow G_L \times G_R$, where the full non-abelian gauge symmetry is restored only when we identify $G_L = G_R$. However, if we break the left gauge group down to $G_L \rightarrow U(1)^{N^2}$, and take the right gauge group to $G_R \rightarrow U(N)$, then we are left with a semi-abelian theory that is gauge invariant in the preffered $U(1)$-directions. We call this manifestly color-dual theory, \textit{semi-abelian Yang-Mills},
\begin{eBox}
\be
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\ee
\end{eBox}
where,
\be
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\ee
\be
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\ee
Here we can think of the abelian vector as a background field that sources on-shell $A_\mu$ currents. Indeed, as we describe in the Appendix \ref{}, semi-abelian Yang-Mills theory is at the heart of YZ-theory \cite{}, J-theory \cite{}, and self-dual Yang-Mills \cite{}. We can use the Ward identity of \ref{eq:GIrelA} to check that the amplitudes are indeed gauge invariant with the abelian polarizations are taken on-shell, $\bar{\varepsilon} \rightarrow k$.

@@

As we have seen, it is possible to coerce tree numerators into 1-loop numerators at any multiplicity for pions.
However, these methods cannot be reapplied to generate higher-loop numerators, which begs the question ``is it possible to find globally color-dual numerators beyond 1-loop''.
We will tackle this question for N$^2$LO 2-to-2 scattering in NLSM before moving on to the same process in YM.

Ignoring tadpoles and BEL graphs, there are 14 cubic 4pt 2-loop graphs (see later picture) which are related to each other via 21 Jacobi relations.
As mentioned in (review section), color-kinematics duality ensures that the numerator of every graph can be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot products of momenta.
The basis of Mandelstam invariants for both of these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram scales as $k^{12}$.
At this mass dimension there are 8,008 independent dot products of momenta.
With two basis graphs there is a total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is not suited to equations with sixteen thousand parameters so a different method is desirable.
One possibility is to use a solver designed for integral reduction.
However, these solvers typically assume that the matrix has a very low density which is not the case for equations coming from NLSM which typically have a density of a few percent.
While the equations in YM naturally separate into different ``helicity'' sectors (or $e_i \cdot e_j$ and $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such separation of sectors in a scalar theory where any product of momenta can interfere with any other.
The NLSM constraint equations were solved using a custom solver, \texttt{FiniteFieldSolve}, designed to solve large linear systems of arbitrary density.
\texttt{FiniteFieldSolve} will be released publicly shortly.}
Imposing all of the Jacobi relations eliminates all but 4,473 of the original free parameters.
Further imposing graph symmetries reduces the number of free parameters to 1,243.

@@

At this point color-kinematics duality has been satisfied, so the double copy will proceed without issue.
However, the plethora of generalized gauge freedom parameters leaves the option for enforcing additional aesthetic constraints.
$\mathcal{N}=4$ SYM provides several hints for further conditions to impose on the ansatz.
For example, for maximally supersymmetric gauge theory it is possible to enforce the no triangle hypothesis, manifest loop power counting, and, at least for 4pt and up to four loops, strip off a factor of $st A_\text{tree}$ from the integrand (I'm not sure I know all of the papers to cite here).
For the NLSM integrand it is desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the 1-loop numerator since this at least manifests antisymmetry for the vertices involving external legs.
However, a more fruitful direction is to match onto the one known theory with pion power counting that manifests color-kinematics duality to all loop order, namely, Zakharov-Mikhailov theory \cite{Zakharov:1973pp, Cheung:2022mix}.

Why don't you just set all of the generalized gauge freedom parameters to zero?
You can learn useful things by not doing this (see N=4), this allows us to introduce more physics into the problem, and, unsurprisingly, setting them to zero produces more a more complex answer with more unwieldy rational numbers than using physics.

ZM theory is governed by the Lagrangian
\be
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2 + g f^{abc} \varphi^a \varepsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c) .
\ee
The color-stripped Feynman rule for the vertex is simply $V(k_1, k_2, k_3) \propto \varepsilon_{\mu\nu}k_1^\mu k_2^\nu$ where off-shell color-kinematics duality to all orders in perturbation theory simply follows from the Schouten identity.
Since the theory only has a cubic vertex and the theory has manifest off-shell color-kinematics duality, it is trivial to read off the color-dual numerator for any graph.
From the presence of the Levi-Civita tensor $\varepsilon^{\mu\nu}$, the theory clearly resides in two spacetime dimensions where scattering is notoriously plagued by infrared regulation issues.
Every on-shell particle is either a right mover, with momentum proportional to $k_R^\mu \equiv (1,1)$, or a left mover, with momentum proportional to $k_L^\mu \equiv (1,-1)$.
On-shell ZM amplitudes naturally divide into sectors corresponding to the configuration of left and right movers, where the scattering in many sectors is rather subtle.
Only \emph{ordered} amplitudes can be defined unambiguously from the Feynman rules.
At 4pt, only the alternating sector is free of subtleties and the amplitude for this process vanishes.
In equations, $A[LRLR]=0$ where $R$ corresponds to a right mover and $L$ corresponds to a left mover.
In every other configuration of right and left movers, such as $A[RRRR]$, one of the internal propagators is incidentally on-shell since $k_R^2=k_L^2=0$.

@@

Proceeding to the next-to-maximal cuts, we continue to avoid
pathological diagrams including the newly-appearing type discussed in
\cref{eq:other-path}. After discarding these types of cuts, we are
left with 8 \ace{(maybe a few more to account for dangling trees)} \jm{How about we just say that there are 8 N$^1$ cuts ignoring dangling trees?}
unique cut topologies, shown in \cref{fig:ym-nmc}, to check.  Seven of
these next-to-maximal cuts are consistent with the kinematic Jacobi
relations and symmetries.  Interestingly, the ``kissing triangles''
cut,
\begin{equation*}
   \NMCutC \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied.
\begin{figure}
\centering
 { \NMCutH, \NMCutG,  \NMCutD,\NMCutB, 
 \\
 \NMCutE, \NMCutF, \NMCutA, \NMCutC,  }
  \caption{The 8 non-pathological next-to-maximal-cut diagrams.}
  \label{fig:ym-nmc}
\end{figure}

@@

\section{Conclusions and Outlook}
In summary, we have found that the two-loop integrand for Yang-Mills cannot be constructed from a color-dual numerator that is a polynomial in the kinematic variables. Thus, if a color-dual representation exists at all-loop order for Yang-Mills, it must be a rational function of kinematics.

This seems to be a rather unnatural constraint on color-dual theories from the perspective of point-like quantum field theories. After all, introducing operators that produce rational functions of kinematics must be non-local. Thus, in order to identify valid color-dual representation of Yang-Mills beyond tree-level, we must relax the constraint of \textit{off-shell locality}, and realize locality only in the limit of \textit{on-shell} kinematics. Below we provide an overview of what one might expect for such a construction, and provide some simple examples at tree-level that manifest this structure. 

@@

\appendix
\section{Catalog of color-dual theories}
\label{sec:CKLagrangians}
\paragraph{Self-Dual Yang-Mills} The first example in the literature \citepls{} of color-dual Feynman rules is that of Yang-Mills theory in the self-dual sector (SDYM):
\be
\mathcal{L}^{\text{YM}} = -\frac{1}{4}F_{\mu\nu}F^{\mu\nu}
\ee
To obtain the Lagrangian for self-dual Yang-Mills, we applyy the self-dual condition that equates the field strength to its dual 2-form:
\be\label{SDcon}
F_{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}F^{\rho\sigma}
\ee
where $F_{\mu\nu} = \partial_\mu A_\nu -\partial_\nu A_\mu + g[A_\mu ,A_\nu]$ is the non-abelian $SU(N)$ field strength. Solutions to the self-dual condition automatically satisfy the Yang-Mills equation of motion due to the Bianchi identity:
\be
D^\mu F^{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}D^\mu F^{\rho\sigma} = 0
\ee
Solutions to \cref{SDcon} correspond to instanton configurations that are probed by the gauge field path integral in the full non-perturbative description of Yang-Mills. In this paper, we will focus on the perturbative aspects of this constraint. Transforming from Cartesian coordinates to light cone coordinantes, $(t,x,y,z) \rightarrow (u,v,w,\bar{w})$, which take the following definition:
\be
u = t+x,\quad v=t-x, \quad w = y+iz
\ee
which gives the following expression for the invariant line element:
\be
ds^2 = 2(dudv - dwd\bar{w}) = dt^2 - d\vec{x}\cdot d\vec{x}
\ee
Applying this coordinate transformation to the self-dual condition of \cref{SDcon}, we obtain the following three constraint equations from the original three cartesian coordinate expressions:
\begin{align} \label{SDLightCone1}
F_{uw} &=0\,,
\\
 \label{SDLightCone2}
 F_{uv} &= F_{w\bar{w}}\,,
\\
 \label{SDLightCone3}
 F_{v\bar{w}} &= 0\,,
\end{align}
Thus, in light-cone gauge, $A_u=0$, we obtain the follow set of equations that are one-to-one with the self-dual condition:
\be
A_u=0 \text{ and } (\ref{SDLightCone1}) \quad\Rightarrow \quad A_w=0
\ee
The \cref{SDLightCone2} then implies the following that $A_v$ and $A_{\bar{w}}$ depend on a single degree of freedom
\be
A_w=0\,,\,A_u=0 \text{ and } (\ref{SDLightCone2}) \quad\Rightarrow \quad \partial_u A_v = \partial_w A_{\bar{w}}
\ee
This constraint is  satisfied with the following definition of the gauge field in terms of a single scalar degree of freedom which we call $\Psi$:
\be\label{SDDef}
A_v = \frac{1}{2} \partial_w \Psi, \qquad A_{\bar{w}} = \frac{1}{2} \partial_u \Psi,
\ee
Using this definition on the final \jm{first?} self-dual constraint equation thus yields the following equation of motion for the dynamical scalar field:
\be\label{SDEOM}
(\ref{SDLightCone1}) \text{ and } (\ref{SDDef}) \quad  \Rightarrow\quad \Box \Psi + ig [\partial_u \Psi, \partial_w \Psi ]=0
\ee
Adding back in the anti-holomorphic field $\bar{\Psi}$ as a Lagrange multiplier, we obtain the following interaction Lagrangian for self-dual Yang-Mills theory:
\begin{eBox}
\be
\mathcal{L}^{\text{SDYM}} = (\partial \bar{\Psi})(\partial \Psi) -i g \bar{\Psi} [\partial_u \Psi, \partial_w \Psi ]
\ee
\end{eBox}
The color-ordered ordered cubic Feynman rule for this theory can be immediately read off as follows:
\be
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =\mathcal{X}(k_1,k_2) \equiv k_{u,1}k_{w,2} - k_{w,1}k_{u,2}
\ee
In this form, the Lagrangian and Feynman rules are manifestly color-dual \textit{off-shell}. In this form, we can see the effect \cref{SDcon} had on the Feynman rules of the full Yang-Mills Lagrangian: it simply decoupled the anti-MHV three-point vertex from the theory, leaving only the MHV three-point vertex. Before proceeding, it is important to note that the mass-dimension of the theory appears to differ from that of Yang-Mills -- we comment on this in detail in our appendix on spinor-helicity variables and conventions.

@@

Next we will briefly describe a 2D scalar theory, Zakharov-Mikhailov (ZM) theory, whose equations of motion bear a striking resemblance to the SDYM equatioi of motion given above in \cref{SDEOM}.

@@

\paragraph{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\be
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\ee
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\be
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\ee
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\be
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\ee
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian
\begin{eBox}
\be
\mathcal{L}^{\text{J}} =\bar{J}^\mu \Box J_\mu +g f^{abc} \bar{J}^\mu J^\nu \partial_\nu J^\mu .
\ee
\end{eBox}
By calculating an off-shell 4pt correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
Kinematic algebra is diff algebra (as usual).
How to extract physical states... (longitudinal J, weird J bar, restores broken Bose).

@@

While the color-dual nature of ZM theory is rather elegant, the loop-level construction introduces complications when applying dimensional regularization. Similar to the difficulties of renormalizing chiral fermions \citepls{ 'tHooft}, there is ambiguity in promoting the integrands to formal $D$-dimensional expressions. As such, in our two-loop bootstrap of NLSM we will only use the ZM integrands as a mechanism for reducing the residual gauge freedom in our final solution -- rather than a full $D$-dimensional integrand to which we match off-shell in dim reg.

\paragraph{Wess-Zumino-Witten theory -- James}
I want to write this down or I'm going to completely forget it

Should probably also mention that BAS has CK at Lagrangian level (duh).

\paragraph{Cubic 2-form theory -- James}

Cubic 2-form theory also has trivial CK.
If you change the propagator (this trick is a theme) then you get YM at least classically.
This is the nonsense ``covariant'' DC.

\paragraph{Chern-Simons theory -- James}
We won't use this but it's useful to treat all of the CK theories in one place
Cite Henrik Maor
Again you place some constraint on the field strength and use Lorenz gauge

\paragraph{Generalities of known color-dual theories -- James}
Should this section maybe go first?
\item This paper kills all ``simple'' color-dual, loop-level Lagrangians for YM but it remains possible that one exists for NLSM.
\item There's also CS but the amplitudes are all zero (and marginally 10D N=1 SYM)
\item All of the theories described above have diff as kinematic algebra, SU(N) for BAS, or rigid/spin Lorenz for cubic 2-form theory
\item All of the theories choose a simple gauge like Lorenz or lightcone
\item All of the theories place some constraint on the field strength directly and get their equations of motion directly from that constraint, not the YM EOM.
\item (Almost all of the theories (aside from NLSM) have trivial amplitudes at least at tree level)
\item This means that almost all of the theories only have CK at the classical level, which can sometimes be fudged into 1-loop as we'll now discuss.
\item For off-shell CK, the kinetic term is relatively unimportant so long as the poles are correctly/consistently identified (what do you mean by s,t,u channel poles).  This might screw up locality in the squared theory.
\section{Spinor-helicity and conventions}
Here we apply the same conventions of Ref.~\cite{jjmcTASI2014}, which we quote now. For massless momenta, $k_a$ and $k_b$, we have 
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]\,,

>>>>

We can achieve this construction by splitting the gauge group of the full Yang-Mills action from $G \rightarrow G_L \times G_R$, where the full non-abelian gauge symmetry is restored only when we identify $G_L = G_R$. However, if we break the left gauge group down to $G_L \rightarrow U(1)^{N^2}$, and take the right gauge group to $G_R \rightarrow U(N)$, then we are left with a semi-abelian theory that is gauge invariant in the preffered $U(1)$-directions. We call this manifestly color-dual theory, \textit{semi-abelian Yang-Mills},
\begin{eBox}
\be
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\ee
\end{eBox}
where,
\be
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\ee
\be
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c .
\ee
Here we can think of the abelian vector as a background field that sources on-shell $A_\mu$ currents. Indeed, as we describe in the Appendix \ref{}, semi-abelian Yang-Mills theory is at the heart of YZ-theory \cite{}, J-theory \cite{}, and self-dual Yang-Mills \cite{}.
Specifically, semi-Abelian YM is simply a clever reinterpretation of the Lagrangian obtained by integrating an auxiliary field into the J-theory equations of motion.
We can use the Ward identity of \ref{eq:GIrelA} to check that the amplitudes are indeed gauge invariant with the abelian polarizations are taken on-shell, $\bar{\varepsilon} \rightarrow k$.

@@

As we have seen, it is possible to coerce tree numerators into 1-loop numerators at any multiplicity for pions.
However, these methods cannot be reapplied to generate higher-loop numerators, which begs the question ``is it possible to find globally color-dual numerators beyond 1-loop''.
We will tackle this question for 4pt 2-loop scattering in NLSM before moving on to the same process in YM.

Ignoring tadpoles and BEL graphs, there are 14 cubic 4pt 2-loop graphs (see \cref{fig:MaxCuts}) which are related to each other via 21 Jacobi relations.
As mentioned in \cref{sec:bootstrap}, color-kinematics duality ensures that the numerator of every graph can be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot products of momenta.
The basis of Mandelstam invariants for both of these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram scales as $k^{12}$.
At this mass dimension there are 8,008 independent dot products of momenta.
With two basis graphs there is a total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is not suited to equations with sixteen thousand parameters so a different method is desirable.
One possibility is to use a solver designed for integral reduction.
However, these solvers typically assume that the matrix has a very low density which is not the case for equations coming from NLSM which typically have a density of a few percent.
While the equations in YM naturally separate into different ``helicity'' sectors (or $e_i \cdot e_j$ and $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such separation of sectors in a scalar theory where any product of momenta can interfere with any other.
The NLSM constraint equations were solved using a custom solver, \texttt{FiniteFieldSolve}, designed to solve large linear systems of arbitrary density.
\texttt{FiniteFieldSolve} will be released publicly shortly.}
Imposing all of the Jacobi relations eliminates all but 4,473 of the original free parameters.
Further imposing graph symmetries reduces the number of free parameters to 1,243.

@@

At this point color-kinematics duality has been satisfied, so the double copy will proceed without issue.
However, the plethora of generalized gauge freedom parameters leaves the option for enforcing additional aesthetic constraints.
$\mathcal{N}=4$ SYM provides several hints for further conditions to impose on the ansatz.
For example, for maximally supersymmetric gauge theory it is possible to enforce the no triangle hypothesis, manifest loop power counting, and, at least for 4pt and up to four loops, strip off a factor of $st A_\text{tree}$ from the integrand (I'm not sure I know all of the papers to cite here).
For the NLSM integrand it is desirable to make color-kinematics duality as manifest as possible.
One hope would be to factor out some piece of the 1-loop numerator since this at least manifests antisymmetry for the vertices involving external legs.
However, a more fruitful direction is to match onto the one known theory with pion power counting that manifests color-kinematics duality to all loop order, namely, Zakharov-Mikhailov (ZM) theory \cite{Zakharov:1973pp, Cheung:2022mix}.

Why don't you just set all of the generalized gauge freedom parameters to zero?
You can learn useful things by not doing this (see N=4), this allows us to introduce more physics into the problem, and, unsurprisingly, setting them to zero produces a more complex answer with more unwieldy rational numbers than using physics.

ZM theory is governed by the Lagrangian
\be
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2 + g f^{abc} \varphi^a \varepsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c),
\ee
where more details can be found in \cref{sec:CKLagrangians}.
The color-stripped Feynman rule for the vertex is simply $V(k_1, k_2, k_3) \propto \varepsilon_{\mu\nu}k_1^\mu k_2^\nu$ where off-shell color-kinematics duality to all orders in perturbation theory simply follows from the Schouten identity.
Since the theory only has a cubic vertex and the theory has manifest off-shell color-kinematics duality, it is trivial to read off the color-dual numerator for any graph.
From the presence of the Levi-Civita tensor $\varepsilon^{\mu\nu}$, the theory clearly resides in two spacetime dimensions where scattering is notoriously plagued by infrared regulation issues.
Every on-shell particle is either a right mover, with momentum proportional to $k_R^\mu \equiv (1,1)$, or a left mover, with momentum proportional to $k_L^\mu \equiv (1,-1)$.
On-shell ZM amplitudes naturally divide into sectors corresponding to the configuration of left and right movers, where the scattering in many sectors is rather subtle.
Only \emph{ordered} amplitudes can be defined unambiguously from the Feynman rules.
At 4pt, only the alternating sector is free of subtleties and the amplitude for this process vanishes.
In equations, $A[LRLR]=0$ where $R$ corresponds to a right mover and $L$ corresponds to a left mover.
In every other configuration of right and left movers, such as $A[RRRR]$, one of the internal propagators is incidentally on-shell since $k_R^2=k_L^2=0$.

@@

Proceeding to the next-to-maximal cuts, we continue to avoid
pathological diagrams including the newly-appearing type discussed in
\cref{eq:other-path}.
After discarding these types of cuts, there are 8 cut topologies (see \cref{fig:ym-nmc}) ignoring cuts with dangling trees.
Seven of these next-to-maximal cuts are consistent with the kinematic Jacobi
relations and symmetries.  Interestingly, the ``kissing triangles'' or ``bowtie''
cut,
\begin{equation*}
   \NMCutC \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied. \jm{I would just focus on what we do since we do something more general than BDN.}
\begin{figure}
\centering
 { \NMCutH, \NMCutG,  \NMCutD,\NMCutB, 
 \\
 \NMCutE, \NMCutF, \NMCutA, \NMCutC,  }
  \caption{The 8 non-pathological next-to-maximal-cut diagrams.}
  \label{fig:ym-nmc}
\end{figure}

@@

\section{Conclusions and Outlook}
There is overwhelming evidence that YM and NLSM obey color-kinematics duality at loop level even though naive Lagrangians for these theories do not manifest the duality.
At tree level, the duality can be manifested by breaking manifest Bose symmetry at the Lagrangian level, either by explicitly picking different gluon states (as in SDYM) or by picking states that only restore Bose symmetry in the final amplitude (as in NLSM).
All of these theories can be encapsulated in the semi-Abelian YM theory presented in \cref{eq:semiYM} which obscures the Bose symmetry of the external states due to the presence of the two gauge fields $A_\mu$ and $\bar{A}_\mu$.
The kinetic mixing between the two gauge fields permits globally color-dual 1-loop integrands, but special care must be taken so that the integrands produced from semi-Abelian YM correspond to SDYM or NLSM.
Due to the kinetic mixing that stems from the broken Bose symmetry, it is impossible to construct a 2-loop integrand for semi-Abelian YM theory, that is, the theory is 1-loop exact.
Obtaining a globally color-dual integrand beyond 1-loop then requires sacrificing more than just Bose symmetry and in this paper it is argued that off-shell locality must be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in order to obtain color-dual 2-loop integrands.}

\jm{I am intentionally shying away from really firm statements like 'you must abandon locality' because there are always loopholes and we made a lot of other assumptions in fixing our \ansatze{}.}

Kinematic ``numerators'' that are rational functions of the kinematics appear unnatural from the perspective of point-like quantum field theories. After all, introducing operators that produce rational functions of kinematics must be non-local. Thus, in order to identify valid color-dual representation of Yang-Mills beyond tree-level, we must relax the constraint of \textit{off-shell locality}, and realize locality only in the limit of \textit{on-shell} kinematics. Below we provide an overview of what one might expect for such a construction, and provide some simple examples at tree-level that manifest this structure. 

@@

\appendix
\section{Catalog of color-dual theories}
\label{sec:CKLagrangians}

Below we review the four manifestly color-dual theories that appear in this paper, in particular SDYM, two formulations of NLSM, and ZM.
All of these theories are closely tied together by the semi-Abelian YM theory in \cref{eq:semiYM} or in other words the cubic sector of YM.
The manifestly color-dual equations of motion for SDYM and NLSM are derived by picking a gauge and then placing a constraint not on the YM equation of motion or Bianchi identity but rather directly on the field strength itself.
Other theories can be obtained from SDYM and NLSM by judicious modification of the propagator.
In all of these theories, the kinematic algebra is that of (volume preserving) diffeomorphisms or some close variant thereof.
In almost all of these theories, Bose symmetry is broken or obscured at the Lagrangian level.
Because of the broken Bose symmetry (and the inherited kinetic mixing of semi-Abelian YM), these theories cannot produce color-dual numerators at 2-loop or higher.

\jm{I've deleted the boxes around equations.  I actually really like them normally but if we're boxing something in the appendix then it feels like that's a major enough result that it should be in the main body of the paper.}

\paragraph{Self-Dual Yang-Mills} The first example in the literature \citepls{} of color-dual Feynman rules is that of Yang-Mills theory in the self-dual sector (SDYM):
\be
\mathcal{L}^{\text{YM}} = -\frac{1}{4}F_{\mu\nu}F^{\mu\nu}
\ee
To obtain the Lagrangian for self-dual Yang-Mills, we applyy the self-dual condition that equates the field strength to its dual 2-form:
\be\label{SDcon}
F_{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}F^{\rho\sigma}
\ee
where $F_{\mu\nu} = \partial_\mu A_\nu -\partial_\nu A_\mu + g[A_\mu ,A_\nu]$ is the non-abelian $SU(N)$ field strength. Solutions to the self-dual condition automatically satisfy the Yang-Mills equation of motion due to the Bianchi identity:
\be
D^\mu F^{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}D^\mu F^{\rho\sigma} = 0
\ee
Solutions to \cref{SDcon} correspond to instanton configurations that are probed by the gauge field path integral in the full non-perturbative description of Yang-Mills. In this paper, we will focus on the perturbative aspects of this constraint. Transforming from Cartesian coordinates to light cone coordinantes, $(t,x,y,z) \rightarrow (u,v,w,\bar{w})$, which take the following definition:
\be
u = t+x,\quad v=t-x, \quad w = y+iz
\ee
which gives the following expression for the invariant line element:
\be
ds^2 = 2(dudv - dwd\bar{w}) = dt^2 - d\vec{x}\cdot d\vec{x}
\ee
Applying this coordinate transformation to the self-dual condition of \cref{SDcon}, we obtain the following three constraint equations from the original three cartesian coordinate expressions:
\begin{align} \label{SDLightCone1}
F_{uw} &=0\,,
\\
 \label{SDLightCone2}
 F_{uv} &= F_{w\bar{w}}\,,
\\
 \label{SDLightCone3}
 F_{v\bar{w}} &= 0\,,
\end{align}
Thus, in light-cone gauge, $A_u=0$, we obtain the follow set of equations that are one-to-one with the self-dual condition:
\be
A_u=0 \text{ and } (\ref{SDLightCone1}) \quad\Rightarrow \quad A_w=0
\ee
The \cref{SDLightCone2} then implies the following that $A_v$ and $A_{\bar{w}}$ depend on a single degree of freedom
\be
A_w=0\,,\,A_u=0 \text{ and } (\ref{SDLightCone2}) \quad\Rightarrow \quad \partial_u A_v = \partial_w A_{\bar{w}}
\ee
This constraint is  satisfied with the following definition of the gauge field in terms of a single scalar degree of freedom which we call $\Psi$:
\be\label{SDDef}
A_v = \frac{1}{2} \partial_w \Psi, \qquad A_{\bar{w}} = \frac{1}{2} \partial_u \Psi,
\ee
Using this definition on the final \jm{first?} self-dual constraint equation thus yields the following equation of motion for the dynamical scalar field:
\be\label{SDEOM}
(\ref{SDLightCone1}) \text{ and } (\ref{SDDef}) \quad  \Rightarrow\quad \Box \Psi + ig [\partial_u \Psi, \partial_w \Psi ]=0
\ee
Adding back in the anti-holomorphic field $\bar{\Psi}$ as a Lagrange multiplier, we obtain the following interaction Lagrangian for self-dual Yang-Mills theory:
\be
\mathcal{L}^{\text{SDYM}} = (\partial \bar{\Psi})(\partial \Psi) -i g \bar{\Psi} [\partial_u \Psi, \partial_w \Psi ]
\ee
The color-ordered ordered cubic Feynman rule for this theory can be immediately read off as follows:
\be
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =\mathcal{X}(k_1,k_2) \equiv k_{u,1}k_{w,2} - k_{w,1}k_{u,2}
\ee
In this form, the Lagrangian and Feynman rules are manifestly color-dual \textit{off-shell}. In this form, we can see the effect \cref{SDcon} had on the Feynman rules of the full Yang-Mills Lagrangian: it simply decoupled the anti-MHV three-point vertex from the theory, leaving only the MHV three-point vertex. Before proceeding, it is important to note that the mass-dimension of the theory appears to differ from that of Yang-Mills -- we comment on this in detail in our appendix on spinor-helicity variables and conventions.

@@

@@

\paragraph{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\be
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\ee
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\be
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\ee
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\be
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\ee
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-Abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
In other words, \emph{J-theory and semi-Abelian YM are identical}.
By calculating an off-shell 4pt correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
As might be expected, the kinematic algebra is the diffeomorphism algebra.
For details about external states and how to extract NLSM (a scalar theory) from semi-YM (a vector theory) we refer the reader to \cite{Cheung:2021zvb}.
Note that the different external particles break manifest Bose symmetry.

@@

While the color-dual nature of ZM theory is rather elegant, the loop-level construction introduces complications when applying dimensional regularization. Similar to the difficulties of renormalizing chiral fermions \citepls{ 'tHooft}, there is ambiguity in promoting the integrands to formal $D$-dimensional expressions. As such, in our two-loop bootstrap of NLSM we will only use the ZM integrands as a mechanism for reducing the residual gauge freedom in our final solution -- rather than a full $D$-dimensional integrand to which we match off-shell in dimensional regularization.

\section{Spinor-helicity and conventions}
\jm{I would be happy to say that we use a mostly minus metric and the spinor helicity conventions of \cite{jjmcTASI2014} in the introduction and omit this appendix.}
